Semantic Prediction in {R}ussian

### 2020-04-11 paper.tex
return 'russian' to the introduction
related works reworking

Our analysis and gold-standard is a cloze test conducted by Laurinavichyute et al \cite{laurinavichyute} from an eye-tracking study. We take the results of the cloze task as a proxy for the underlying probability distribution of next-word continuations of partial sentences.
The LM based on these answers we will call "human-like". So the uniqueness of the research is that we can compare artificial language models with the model which approximates real human expectations about the next word for a given sequence.

@@

As Kuperberg and Jaeger claim in their “What do we mean by prediction
in language comprehension?” \cite{kuperberg-jaeger}, the reaction time is in direct ratio with the predictability of the word: the more predictable the word is the faster is the reaction time. Moreover, predictability of a word or a context defines fixation time in eye-movement studies as a result of the language comprehension process. This implies that language comprehension must be predictive. The authors also state that as the previous context expands, the predictability of the next word increases leading to - in cloze tests - higher accuracy of predicting the next word, and - in eye-movement experiments - to shorter fixation duration.

Several scientific papers propose different algorithms for cloze answer generation. In \cite{zhou-cloze} the authors state the importance of next word prediction in language modeling and its potential contribution to OCR and handwriting recognition. The authors enhance existing models with ELMo and BERT language models and train on the CLOTH dataset of cloze tests. Expectedly, BERT models show the highest performance (0.86 and 0.83 accuracy scores on test dataset for BERT Large and BERT Base respectively), as this model was initially trained to recover masked tokens in text. At the same time, ELMo model's poor performance could be due to the lack of parameter tuning and the fact that ELMo was trained for the next sequence word prediction.

@@

To evaluate each model’s predictions, we took the first 30 most probable words. Each probability was renormalized by dividing originally computed word-wise probability by the sum of probabilities of the first 30 words. This way the sum of probabilities the selected words would equal 1.

@@

Firstly, we converted all of the modified Pymorphy tags into 4 general sets: 'ADJ', 'ADVB', 'NUMR' were generalized to 'MOD'; 'INFN' and 'PRED' - to 'VERB'; 'NPRO' to 'NOUN'; 'PREP', 'PRCL', 'CONJ' and 'INTJ' to 'FUNC' for each context. Then, we have counted probabilities of these tags in the same manner as the usual parts of speech.

We noticed that the generalized probabilities were overall higher than with modified Pymorphy tags.

@@

Then, MiniBatchKmeans (a modified version of the K-means algorithm,
that uses mini-batches to reduce computation time, while at the same
time trying to optimize the same goal function \cite{bejar}) algorithm in sklearn was used to find “cluster centers”, or mean semantic vector for one model for each context. And at last, we found cosine similarity (also with sklearn library) for each pair of semantic vectors for each pair of models.

@@

In Table 3, contexts with minimum variance in filling
are listed. It is worth emphasizing that in all the above cases there was no variance in respondents’ answers, i.e. all respondents gave the same one answer for these contexts. What is more, this predicted word was the original word from the corpus.

@@

To measure model predictions on the semantic level, the cosine similarity between each context’s predicted words centroid vector was found. The number of words was selected dynamically for each context by maximizing vector significance with the minimum words. Figure 6 reflects the results.
\begin{figure}
\caption{Cosine Similarity between each pair of models .}
\label{fig:ks-lexical}
\centering
\includegraphics[width=0.5\textwidth]{figures/pdf/cosine-similarity-lexical.pdf}
\end{figure}

>>>>

Our analysis and gold-standard is a Russian cloze test conducted by Laurinavichyute et al \cite{laurinavichyute} from an eye-tracking study. We take the results of the cloze task as a proxy for the underlying probability distribution of next-word continuations of partial sentences.
The LM based on these answers we will call "human-like". So the uniqueness of the research is that we can compare artificial language models with the model which approximates real human expectations about the next word for a given sequence.

@@

As Kuperberg and Jaeger claim in their “What do we mean by prediction in language comprehension?” \cite{kuperberg-jaeger}, the reaction time is in direct proportion with the predictability of the word: the more predictable the word is the faster is the reaction. Moreover, predictability of a word or a context defines fixation time in eye-movement studies as a result of the language comprehension process. This implies that language comprehension must be predictive. The authors also state that as the previous context expands, the predictability of the next word increases leading to - in cloze tests - higher accuracy of predicting the next word, and - in eye-movement experiments - to shorter fixation duration.

The literature contains several different algorithms for cloze answer generation. In \cite{zhou-cloze} the authors state the importance of next word prediction in language modeling and its potential contribution to OCR and handwriting recognition. The authors enhance existing models with ELMo and BERT language models and train on the CLOTH dataset of cloze tests. BERT models show the highest performance (0.86 and 0.83 accuracy scores on test dataset for BERT Large and BERT Base respectively), as this model was initially trained to recover masked tokens in text. At the same time, the ELMo model's poor performance could be due to the lack of parameter tuning and the fact that ELMo was trained for the next sequence word prediction.

@@

To evaluate each model’s predictions, we took the first 30 most probable words. Each probability was renormalized by dividing originally computed word-wise probability by the sum of probabilities of the first 30 words. This way the sum of probabilities the selected words would equal to 1.

@@

Firstly, we converted all of the modified Pymorphy tags into 4 general sets: 'ADJ', 'ADVB', 'NUMR' were generalized to 'MOD'; 'INFN' and 'PRED' - to 'VERB'; 'NPRO' to 'NOUN' (Object); 'PREP', 'PRCL', 'CONJ' and 'INTJ' to 'FUNC' for each context. Then, we have counted probabilities of these tags in the same manner as the usual parts of speech.

We noticed that the generalized probabilities were overall higher than with modified Pymorphy tags - further we will refer to it as OMVF (object-modifier-verb-functional).

@@

Then, MiniBatchKmeans (a modified version of the K-means algorithm, that uses mini-batches to reduce computation time, while at the same time trying to optimize the same goal function \cite{bejar}) algorithm in sklearn was used to find “cluster centers”, or mean semantic vector for one model for each context. And at last, we computed cosine similarity (also with sklearn library) for each pair of semantic vectors for each pair of models.

@@

In Table 3, contexts with minimum variance in filling
are listed. It is worth emphasizing that in all these cases there was no variance in respondents’ answers, i.e. all respondents gave the same one answer for these contexts. What is more, this predicted word was the original word from the corpus.

@@

To measure model predictions on the semantic level, the cosine similarity between each context’s predicted words centroid vector was found. The number of words was selected dynamically for each context by maximizing vector significance with the minimum words. Figure 6 reflects the results.
\begin{figure}
\caption{Cosine Similarity between each pair of models .}
\label{fig:ks-lexical-pairs}
\centering
\includegraphics[width=0.5\textwidth]{figures/pdf/cosine-similarity-lexical.pdf}
\end{figure}
